The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow

### 2016-09-30 paper/thingap_mri_paper.tex
thingap appendices

\section{Equations}

@@

The partial differential equations that comprise Equations \ref{eq:ordere} to \ref{eq:ordere3} are solved in succession. The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension. Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations. We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension. We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep). We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g. \citei{Boyd:2001aa}). The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary layers well on a 128-point grid.

@@

>>>>

\section{Equations}\label{sec:equations}

@@

The partial differential equations that comprise Equations \ref{eq:ordere} to \ref{eq:ordere3} are solved in succession. The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension. Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations. We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension. We solve the radial component of each equation using the open source pseudospectral code Dedalus. We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g. \citei{Boyd:2001aa}). The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary layers well on a 128-point grid.

@@

%************* APPENDICES ************************%

\clearpage
\appendix

\section{A. Detailed Equations}\label{app:matrices}

Here we detail the perturbation analysis described in Section \ref{sec:equations}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where

\beq
\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
\eeq

\beq
\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
\eeq

and the constituent matrices are defined as

\beq
\mathcal{D} = \left[\begin{matrix}
\nabla^2 & 0 & 0 & 0 \\
0 & 1& 0 & 0 \\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1 \\
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_0 = \left[\begin{matrix}
-\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
0 & -\frac{1}{\reye}\partial_x^2 & 0 &0 \\
0 & 0 & -\frac{1}{\reym}\partial_x^2 & 0 \\
0 & 0 & 0 & -\frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
\eeq

\beq
\mathcal{L}_1 = \left[\begin{matrix}
0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
0 & -1 & q\Omega_0 & 0 \\ \end{matrix}\right] 
\eeq

\beq
\mathcal{L}_2 = \left[\begin{matrix}
-2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
0 & -\frac{1}{\reye} & 0 & 0 \\
0 & 0 & -\frac{1}{\reym} & 0 \\
0 & 0 & 0 & -\frac{1}{\reym} \\ \end{matrix}\right]
\eeq

\beq
\mathcal{L}_3 = \left[\begin{matrix}
0 & 0 & -\frac{2}{\beta} & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\ \end{matrix} \right]
\eeq

\beq
\mathcal{L}_4 = \left[\begin{matrix}
-\frac{1}{\reye} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\ \end{matrix}\right] 
\eeq

\beq
\mathcal{G} = \left[\begin{matrix}
0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
0 & 0 & 0 & -\frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
0 & -1 & 0 & 0 \\ \end{matrix} \right]
\eeq

\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
\beq
\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
\eeq

where

\beq
N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
\eeq

\beq
N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
\eeq

\beq
N_2^{(A)} = - J(A_1, \Psi_1) 
\eeq

\beq
N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
\eeq

and

\beq
\begin{split}
N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
\end{split}
\eeq

\beq
N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
\eeq

\beq
N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
\eeq

\beq
N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
\eeq

$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become

\beq
\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
\eeq

and

\beq
\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
\eeq

The second order nonlinear terms are \\

\beq
\begin{split}
N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
\end{split}
\eeq

\beq
\begin{split}
N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
\end{split}
\eeq

\beq
N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
\eeq

\beq
N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
\eeq

\beq
\begin{split}
N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
\end{split}
\eeq

\beq
N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
\eeq

\beq
N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
\eeq

\beq
N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
\eeq \\

and the third order nonlinear terms become \\

\beq
\begin{split}
N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
\end{split}
\eeq

\beq
\begin{split}
N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
\end{split}
\eeq

%Correct -- checked 11/14/15
\beq
\begin{split}
N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
 + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
 & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
\end{split}
\eeq

\beq
\begin{split}
N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
\end{split}
\eeq
